---
title: "ROMOPOmics - An OMOPOmics R package"
date: June 16, 2020
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dbplyr)
library(data.table)
library(rmarkdown)
library(knitr)
library(kableExtra)
#library(here)
library(rprojroot)
library(RSQLite)
library(ROMOPOmics)
knitr::opts_chunk$set(echo = TRUE)

dirs          <- list()
#Something wierd about how here() works within and without a markdown; check for interactivity.
#dirs$base <- ifelse(interactive(),
#                    file.path(here(),
#                    file.path(here(),"demo"))
#dirs$base     <- here()
dirs$base     <- file.path(find_root(criterion = criteria$is_r_package))
dirs$data     <- file.path(dirs$base,"data")
dirs$demo     <- file.path(dirs$base,"demo")
dirs$masks    <- file.path(dirs$base,"masks")

#Files.
dm_file     <- system.file("extdata","OMOP_CDM_v6_0_custom.csv",package="ROMOPOmics",mustWork = TRUE)
tcga_files  <- dir(dirs$data,full.names = TRUE)
names(tcga_files) <- gsub(".csv","",basename(tcga_files))

knitr::opts_chunk$set(echo = FALSE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
