---
title: "Nof1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nof1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup,message=F}
library(ROMOPOmics)
library(tidyverse)
```

# Creating a database of 1

Let's say we have a cancer patient Sally that has undergone treatment and genomic testing. If we have that information available to us, we can generate her own database of clinical and 'omics data.

# Data Model

```{r loadDataModel,echo=TRUE}

dm <- loadDataModel(as_table_list = FALSE)

```

# Input Data

We need to have a matrix of information for Sally

```{r data_generation}

df <- 
data.frame(
  "patient" = "Sally",
  "tumor_type" = "glioblastoma",
  "cancer_report" = "MSK Impact report",
  "specimen" = "brain",
  "mutation_1" = "ARID1A",
  "mutation_2" = "ARID1B"
  )

df %>% write_csv("../data/Nof1.csv")

df

```

# Mask Data

We need to have a mask in order to map to OMOP's standard concepts, fields, and tables

```{r mask_generation}

mask_df <- 
data.frame(
  "alias" = 
    colnames(df),
  "example1" = 
    df[1,] %>% unlist() %>% levels(),
  "field" = 
    c("person_id","specimen_source_value","note_title","anatomic_site_source_value","gene_name","gene_name"),
  "table" = 
    c("person","specimen","note","specimen","variation","variation"),
  "field_idx" = 
    c(NA,NA,NA,NA,1,2),
  "set_value" =
    c(NA,NA,NA,NA,"copy number variation","SNP")
)

mask_df %>% write_csv("../data/Nof1_mask.csv")

mask_df

```

# Load Mask Data

```{r load_masks}

msks <- loadModelMasks(mask_files = "../data/")

```

# Translate into Common Data Model

```{r translate}

nof1data <- list.files('../data/',"Nof1.csv")

omop_inputs <- 
  readInputFile(
    input_file = "../data/Nof1.csv",
    mask_table = msks$Nof1,
    data_model = dm
    )

omop_inputs

```

# Make database input tables

```{r db}

db_inputs   <- combineInputTables(input_table_list = omop_inputs)

```
